---
title: Basic of `{ggplot2}`
---

### Getting Started

1. **Load your library and data**
```{r}
#| message: false
library(tidyverse)

data <- read.csv("../data/Student_Mental_health.csv")

```

2. **Part of** `{ggplot2}`

`{ggplot2}` has several relavant functions:

- `ggplot()` with the **arguments**:

    - `data`
    - `aes`
    - `geom`
    - `stats`
    - `facets`
    - `scales`
    - `theme`
    - `labs`
    - and a lot more.

The key essential arguments are:  `data`, `aes`, `geom`. Then, you can customize your plot.

Find the pre-processed data here

```{r}
#| code-fold: true
data <- data %>%
          mutate(gender = factor(Choose.your.gender))

```

The **basic**, here you see the data the `ggplot` and the **aesthetics**
```{r}
data %>%
  ggplot(aes(x= gender)) 
```

Add a **geometric objects** (use `+`insted of `%>%`)

    -   `geom_point()` → Scatter plot
	-   `geom_line()` → Line plot
	-   `geom_bar()` → Bar chart
	-   `geom_histogram()` → Histogram
	-   `geom_boxplot()` → Box plot
	-   `geom_density()` → Density plot

```{r}
data %>%
  ggplot(aes(x= gender)) +
  geom_bar()
```

Modify the **axis**, add a title.

```{r}
data %>%
  ggplot(aes(x= gender)) +
  geom_bar() +
  labs(
    x     = "Gender classification",
    y     = "Number of participants",
    title = "Distribution of Gender"
  )

```

You can also change the background using `theme_bw()`. The `theme()` function is highly customizable, allowing you to adjust various elements of your plot. `theme_bw()` is one of the predefined themes that provides a clean, minimalistic look with a white background and black grid lines.

```{r}
data %>%
  ggplot(aes(x= gender)) +
  geom_bar() +
  labs(
    x     = "Gender classification",
    y     = "Number of participants",
    title = "Distribution of Gender Frequency "
  ) +
  theme_bw()
```